---
title: "Exploratory - _C. difficile_ by time"
author: "Nicholas Lesniak"
date: "2/9/2017"
output: html_document
---

```{r setup environment, include=F}
knitr::opts_chunk$set(echo = F, warning = F, message = F)

pack_used <- c('tidyr','dplyr','ggplot2')
for (dep in pack_used){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), repos = 'http://cran.us.r-project.org', 
                     quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}

meta_file   <- '../data/process/abx_cdiff_metadata_clean.txt'
shared_file <- '../data/mothur/abx_time.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.an.unique_list.0.03.subsample.shared'
tax_file <- '../data/mothur/abx_time.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.an.unique_list.0.03.cons.taxonomy'
tax_function <- '../code/sum_otu_by_taxa.R'

meta_file   <- read.table(meta_file, sep = '\t', header = T, stringsAsFactors = F)
shared_file <- read.table(shared_file, sep = '\t', header = T, row.names = 'Group')
source(tax_function)

cdiff_otus <- read.table('../data/mothur/abx_time.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.an.unique_list.0.03.cons.taxonomy', sep = '\t', header = T)
cdiff_otus <- as.character(cdiff_otus$OTU[grep('Clostridium_XI', cdiff_otus$Taxonomy)])

shared_file <- select(shared_file, -label, -numOtus, -one_of(cdiff_otus))

```


```{r shared setup, include=F}

#OTU_total <- if(summary(apply(shared_file, 1, sum))['Min.'] ==
#   summary(apply(shared_file, 1, sum))['Max.']){
#  as.numeric(summary(apply(shared_file, 1, sum))['Mean'])
#}


#counts[which(counts <= 1000)]
#counts[1294]
#shared_file[1294,which(test[1294,] > 0)]
#test[1294,which(test[1294,] > 0)]

# remove OTUs not present in at least percentage of samples
otu_presence_cutoff <- 0.05
otu_presence <- ifelse(shared_file > 0, T, F) 
OTUs_present <- colnames(shared_file[ , apply(otu_presence, 2, sum) > 
                                        round(otu_presence_cutoff*nrow(shared_file))])
#relative abundance of OTUs present in >5% of samples
#rel_abund <- shared_file[ ,OTUs_present]/OTU_total*100
counts <- apply(shared_file,1,sum)
rel_abund <- data.frame(apply(shared_file[ ,OTUs_present], 2, function(i){(i/counts)*100}))

# fix meta file 
meta_file$dose <- as.character(meta_file$dose)
meta_file$dose[meta_file$dose == '10mg/kg'] <- 10
meta_file$dose <- as.numeric(meta_file$dose)

meta_file$abx[meta_file$abx == 'vanc '] <- 'vanc'

```

```{r community dynamics}
meta_shared <- right_join(meta_file, mutate(rel_abund, group = rownames(rel_abund)))

phyla_rel_ab <- sum_otu_by_taxa(taxonomy_file = tax_file, otu_df = rel_abund,taxa_level = 'phylum')
meta_phyla <- right_join(meta_file, mutate(phyla_rel_ab, group = rownames(rel_abund)))
taxa <- colnames(select(phyla_rel_ab, -sample))

all_plots <- lapply(unique(meta_phyla$abx),function(antibiotic){
cfu_plot <- meta_phyla %>% 
    filter(day >= 0, abx == antibiotic) %>% 
      ggplot(aes(x = day, y = CFU +1, color = dose)) + 
        scale_y_log10() + 
        scale_colour_gradient(low = '#FFCCCC', high = '#990000', name = 'Antibiotic Dose') + 
        geom_jitter(width = 0.3, alpha = 0.2) + 
        stat_summary(fun.y=median, geom="line", aes(group = cage)) +
        stat_summary(fun.y=median, geom="line", size = 1, color = 'black') +
        labs(x = 'Day', y = ' CFU (log10)', 
             title = paste0('C. Difficile Colonization - ', antibiotic)) +
        stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5), color = 'black') + 
        theme_bw()

community_plot <- meta_phyla %>% 
  filter(day >= 0, abx == antibiotic) %>% 
  gather(taxonomy, abundance, one_of(taxa)) %>% 
  ggplot(aes(x = day, y = abundance, color = dose)) + 
        scale_colour_gradient(low = '#FFCCCC', high = '#990000', name = 'Antibiotic Dose') + 
        facet_grid(taxonomy ~., scales = 'free_y') + 
        geom_jitter(width = 0.3, alpha = 0.2) + 
        stat_summary(fun.y=median, geom="line", aes(group = cage)) +
        labs(x = 'Day', y = ' Relative Abundance', title = paste(antibiotic)) +
        stat_summary(fun.y=median, geom="line", size = 1, color = 'black') +
        stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5), color = 'black') + 
        theme_bw() +
        theme(legend.title=element_blank())
return(list(cfu_plot, community_plot))
})

antibiotics <- c('Clindamycin','Cefaperazone','Streptomycin')
abxs <- c('clinda','cef','strep')
meta_phyla$CFU[is.na(meta_phyla$CFU)] <- 0

for(i in abxs){        
  # single cdiff colonization plot
  ggsave(paste0(i, '_cdiff_single_mouse.png'), width = 5.5, height = 5.5, 
    meta_phyla %>%
      filter(day >= 0, abx == i, cage %in% c(4,104,48), mouse == 1) %>% 
      ggplot(aes(x = day, y = CFU +1)) + 
          scale_y_log10(limits = c(1, 1e9), 
            breaks = scales::trans_breaks("log10", function(x) 10^x),
            labels = scales::trans_format("log10", scales::math_format(10^.x))) + 
          geom_line(lwd = 1.5, color = 'red') + 
          labs(x = 'Day', y = ' CFU (log10)', 
            title = paste0('C. Difficile Colonization - ', antibiotics[abxs == i])) +
          scale_x_continuous(expand = c(0, 0), breaks = seq(0, 10, by = 1)) + 
          theme_bw()
          )
  # single cdiff community plot
  plot <- meta_phyla %>%
    filter(day >= 0, abx == i, cage %in% c(4,104,48), mouse == 1) %>% 
    gather(taxonomy, abundance, one_of(taxa)) %>% 
    ggplot(aes(x = day, y = abundance, color = taxonomy)) + 
      geom_line(lwd = 1.5) + 
      labs(x = 'Day', y = ' Relative Abundance', 
        title = paste0('Community Dynamics - ', antibiotics[abxs == i])) +
      scale_x_continuous(expand = c(0, 0), breaks = seq(0, 10, by = 1)) + 
      scale_y_continuous(expand = c(0, 0), limits = c(0,100)) + 
      theme_bw()

  if(i == 'clinda'){
    ggsave(paste0(i, '_community_single_mouse.png'), width = 5.5, height = 5.5,
      plot + 
      theme(legend.position = c(0.95, 0.5),
        legend.justification = c(1, 0.5), 
        legend.background = element_rect(colour = 'black', fill = "white"),
        legend.title=element_blank()))
  } else {
    ggsave(paste0(i, '_community_single_mouse.png'), width = 5.5, height = 5.5, 
      plot + theme(legend.position = 'none')
      )
  }
}


```

***
***


## None

```{r cfu plot, fig.width=9, fig.height=3, eval=F}
all_plots[[1]][[1]]
```

```{r community plot, fig.width=10, fig.height=9, eval=F}
all_plots[[1]][[2]]
```

## Clindamycin

```{r cfu plot clinda, fig.width=9, fig.height=3}
all_plots[[2]][[1]]
```

```{r community plot clinda, fig.width=10, fig.height=9}
all_plots[[2]][[2]]
```

## Vancomycin

```{r cfu plot vanc, fig.width=9, fig.height=3}
all_plots[[3]][[1]]
```

```{r community plot vanc, fig.width=10, fig.height=9}
all_plots[[3]][[2]]
```

## Ciprofloxacin

```{r cfu plot cipro, fig.width=9, fig.height=3}
all_plots[[4]][[1]]
```

```{r community plot cipro, fig.width=10, fig.height=9}
all_plots[[4]][[2]]
```

## Ampicillin

```{r cfu plot amp, fig.width=9, fig.height=3}
all_plots[[5]][[1]]
```

```{r community plot amp, fig.width=10, fig.height=9}
all_plots[[5]][[2]]
```

## Cefaperazone

```{r cfu plot cef, fig.width=9, fig.height=3}
all_plots[[6]][[1]]
```

```{r community plot cef, fig.width=10, fig.height=9}
all_plots[[6]][[2]]
```

## Metronidazole

```{r cfu plot met, fig.width=9, fig.height=3}
all_plots[[7]][[1]]
```

```{r community plot met, fig.width=10, fig.height=9}
all_plots[[7]][[2]]
```

## Streptomycin

```{r cfu plot strep, fig.width=9, fig.height=3}
all_plots[[8]][[1]]
```

```{r community plot strep, fig.width=10, fig.height=9}
all_plots[[8]][[2]]
```


```{r eval=FALSE}


meta_file %>% 
  filter(day >= 0) %>%
  ggplot(aes(x = day, y = CFU +1, color = dose)) + 
        scale_y_log10() + 
        scale_colour_gradient(low = '#FFCCCC', high = '#990000', limits=c(0, 10)) + 
        facet_grid(abx ~., scales = 'free_y') + 
        geom_jitter(width = 0.3, alpha = 0.2) + 
    #    scale_color_manual(values = c('black', 'red'), breaks=c("FALSE", "TRUE"),
    #                       labels=c("Persistent", "Severe")) +
        stat_summary(fun.y=median, geom="line", aes(group = cage)) +
        labs(x = 'Day', y = ' CFU (log10)', title = 'C. Difficile Colonization') +
        stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5)) + 
        theme_bw() +
        theme(legend.justification=c(1,0), legend.position=c(1,0.25), 
              legend.title=element_blank())


```

## diversity vs colonization

###alpha diversity
####communities colonized to higher levels have lower diversity (alpha)
#### association between cfu and alpha (invsimpson and shannon - NS)
#### cfu vs # of otus (NS)
#### shared otus?
###beta diversity
####highly infected communities are most different than untreated
#### separation between untreated mice and all the highly infected communities (>1e6)
####communities that recover/elimnate cdifficile are more diverse
#### difference in diversity between highly infected 
####more change w/low diversity?
####more change with high cfu?

####need to remove dependence of daily sampling?


```{r diversity}
library(tidyverse)


# reduce samples to no abx or high abx with no delay
hi_lo_no_delay_df <- meta_file %>% 
  filter(delayed == F, (abx %in% c('none','amp','cipro','clinda','metro')) |
    (abx == 'cef' & dose == 0.5) | (abx == 'strep' & dose == 5) |
    (abx == 'vanc' & dose == 0.625)) %>% 
  select(group, CFU, cage, mouse, day, abx, cdiff, preAbx)


alpha_df <- read.table(file = "../data/mothur/abx_time.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.an.unique_list.groups.ave-std.summary", header = T) %>% 
  filter(method == 'ave')

CFU_vs <- function(variable_name, minimum){
  data_frame <- hi_lo_no_delay_df %>% 
    filter(day > 0, cdiff == TRUE) %>% 
    filter(CFU > minimum) %>% 
    select(group, CFU) %>% 
    mutate (CFU = CFU + 0.000000001) %>% 
    left_join(select(alpha_df, group, variable_name))
  plot <- data_frame %>% 
    ggplot(aes(x = get(variable_name), y = CFU)) + 
      geom_point() + 
      scale_y_log10() + 
      geom_smooth(method = 'lm') +
      theme_bw() + 
      labs(x = variable_name, y = 'C.difficile CFU')
  fit_data <- summary(lm(get(variable_name) ~ CFU, data_frame))
  return(list(plot, fit_data))
}

invsimpson_CFU <- CFU_vs('invsimpson', -1)
sobs_CFU <- CFU_vs('sobs', -1)
shannon_CFU <- CFU_vs('shannon', -1)

invsimpson_CFU_wo0 <- CFU_vs('invsimpson', 0)
sobs_CFU_wo0 <- CFU_vs('sobs', 0)
shannon_CFU_wo0 <- CFU_vs('shannon', 0)


```


## Alpha Diversity

```{r invsimpson, fig.width=4, fig.height=3}
invsimpson_CFU[1]
invsimpson_CFU[2]
```

```{r sobs, fig.width=4, fig.height=3}
sobs_CFU[1]
sobs_CFU[2]
```

```{r shannon, fig.width=4, fig.height=3}
shannon_CFU[1]
shannon_CFU[2]
```


```{r invsimpson_no_0, fig.width=4, fig.height=3}
invsimpson_CFU_wo0[1]
invsimpson_CFU_wo0[2]
```

```{r sobs_no_0, fig.width=4, fig.height=3}
sobs_CFU_wo0[1]
sobs_CFU_wo0[2]
```

```{r shannon_no_0, fig.width=4, fig.height=3}
shannon_CFU_wo0[1]
shannon_CFU_wo0[2]
```



```{r high vs low}


otus_df <- read.table(file = "../data/mothur/abx_time.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.an.unique_list.0.03.subsample.shared", header = T)
tax_df <- read.table(file = "../data/mothur/abx_time.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.an.unique_list.0.03.cons.taxonomy", header = T)

#convert abundance to presence (T/F)
otu_presence <- otus_df %>%
  gather(otu, abundance, contains('Otu00')) %>%
  mutate(presence = ifelse(abundance > 0, T, F)) %>%
  select(Group, otu, presence) %>%
  spread(otu, presence)
#subset metadata to samples
presence_df <- metadata %>%
  filter(day > 0, cdiff == TRUE) %>% 
  select(group, CFU, day) %>% rename(Group = group) %>% 
  filter(!between(CFU, 1, 1e5)) %>%
  mutate(colonization = ifelse(CFU < 1, F, T)) %>%
  inner_join(otu_presence)

# merge otu presence with colonization
presence_by_day_df <- presence_df %>%
  select(colonization, day, contains('Otu00')) %>%
  gather(otu, presence, contains('Otu00')) %>%
  group_by(colonization, otu, day) %>%
  summarise_all(funs(mean)) %>%
  ungroup()
presence_by_day_df$otu_labels <- gsub("Otu0*", "", presence_by_day_df$otu)

#test for significant differences between colonized/uncolonized by otu
sig_presence <- c()
for(i in unique(presence_by_day_df$otu)){
  colonized <- presence_by_day_df %>%
    filter(otu == i & colonization == T) %>%
    select(presence)
  uncolonized <- presence_by_day_df %>%
    filter(otu == i & colonization == F) %>%
    select(presence)
  sig_presence <- rbind(sig_presence,
    list(i, as.numeric(wilcox.test(c(colonized$presence), c(uncolonized$presence))$p.value)))
}
significant_otu_list <- unlist(sig_presence[unlist(sig_presence[,2]) < 0.05/nrow(sig_presence), 1])

sig_by_day_df <- presence_by_day_df %>%
  filter(otu %in% significant_otu_list) 

sig_by_day_df %>%
  ggplot(aes(x = otu, y = presence, color = colonization)) + 
    stat_summary(fun.data=median_hilow, position=position_dodge(0.2), fun.args=list(conf.int=0.5)) +
    labs(title = 'Comparison of OTUs present in Colonized/Uncolonized Samples',
      subtitle = '(Median and IQR shown)\nAll OTUs plotted are significant (p < 0.05) after BF multiple comparisons correction ',
      y = 'Portion of Samples with OTU Present', x = 'OTU') + 
    scale_color_manual(labels = c("Uncolonized (C. difficle CFU = 0)", "Colonized (C. difficle CFU >= 1e5)"), values = c("blue", "red")) +
    scale_x_discrete(labels=unique(sig_by_day_df$otu_labels)) + 
    theme_bw() + 
    theme(legend.justification=c(1,1), legend.position=c(0.95,0.95), 
      legend.title = element_blank(),   legend.box.background = element_rect(),
      legend.box.margin = margin(0, 0, 0, 0),
      axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r rel abund}
# by relative abundance
# presence in at least 10%
otus_in_samples <- presence_by_day_df %>% 
  group_by(otu) %>% 
  summarize(max_presence = mean(presence)) %>% 
  filter(max_presence >= 0.1) %>% 
  select(otu)

rel_abun_by_day <- metadata %>%
  filter(day > 0, cdiff == TRUE) %>% 
  select(group, CFU, day) %>% rename(Group = group) %>% 
  filter(!between(CFU, 1, 1e5)) %>%
  mutate(colonization = ifelse(CFU < 1, F, T)) %>%
  inner_join(select(otus_df, Group, contains('Otu00'))) %>% # switch from using 10% presence to median 1% abundance
  gather(otu, abundance, contains('Otu00'))  %>% 
  mutate(rel_abund = abundance / otus_df$numOtus[1])

# at least 1% median abundance in a group
otus_by_median <- rel_abun_by_day %>% 
  group_by(colonization, otu) %>% 
  summarise(median=median(rel_abund)) %>% 
  ungroup() %>% group_by(otu) %>% 
  summarise(median=max(median)) %>% 
  filter(median >= 0.01) %>% 
  select(otu)

rel_abun_df <- filter(rel_abun_by_day, otu %in% otus_by_median$otu)

sig_abundance <- c()
for(i in unique(rel_abun_df$otu)){
  colonized <- rel_abun_df %>%
    filter(otu == i & colonization == T) %>%
    select(rel_abund)
  uncolonized <- rel_abun_df %>%
    filter(otu == i & colonization == F) %>%
    select(rel_abund)
  sig_abundance <- rbind(sig_abundance,
    list(i, as.numeric(wilcox.test(c(colonized$rel_abund), c(uncolonized$rel_abund))$p.value)))
}
sig_relabund_otu_list <- unlist(sig_abundance[unlist(sig_abundance[,2]) < 0.05/nrow(sig_abundance), 1])

sig_relabun_by_day_df <- rel_abun_df %>%
  mutate(otu_labels = gsub("Otu0*", "", rel_abun_df$otu)) %>% 
  filter(otu %in% sig_relabund_otu_list) 
  


sig_relabun_by_day_df %>%
  ggplot(aes(x = otu, y = rel_abund, color = colonization)) + 
    stat_summary(fun.data=median_hilow, position=position_dodge(0.2), fun.args=list(conf.int=0.5)) +
    labs(title = 'Comparison of Relative Abundance in Colonized/Uncolonized Samples',
      subtitle = '(Median and IQR shown)\nAll OTUs plotted are significant (p < 0.05) after BF multiple comparisons correction ',
      y = 'Relative Abundance', x = 'OTU') + 
    scale_color_manual(labels = c("Uncolonized (C. difficle CFU = 0)", "Colonized (C. difficle CFU >= 1e5)"), values = c("blue", "red")) +
    scale_x_discrete(labels=unique(sig_relabun_by_day_df$otu_labels)) + 
    theme_bw() + #coord_cartesian(ylim = c(0, 0.01)) + 
    theme(legend.justification=c(1,1), legend.position=c(0.95,0.95), 
      legend.title = element_blank(),   legend.box.background = element_rect(),
      legend.box.margin = margin(0, 0, 0, 0),
      axis.text.x = element_text(angle = 45, hjust = 1))


```




After talking with Pat (1/17/18)
  Does end point look like initial?
    are most similar the recovered ones?
  Split analysis by abx/dose/days recovered
  Resistance more similar than colonized?
  Start with high dose and 1 day recovery
    then look at how modulating the dose/recovery affects 
    train model with low recovery and test with high recovery
  Show different context of day 0
  compare differences in metro recovery 
    how do susceptibility break points compare?

the metadata file has the following columns
  group
  CFU - ranges 0 to 8.1e8 with 601 NAs 
    (most of NAs are on days when cdiff was not present, so can change to 0 expect for NAs after day 1)
  cage
  mouse
  day - ranges -11 to 10
  abx - 
    amp    cef  cipro clinda  metro   none  strep   vanc 
    405    379     83    190    339      3    362    312 
  dose - 
    0.1     0.3     0.5   0.625       1 10mg/kg       5    NA's 
    304     253     653     112     339     273     136       3 
        dose    abx     cages   mice
        <NA>    none    1       1
        0.5     amp     
        10      cipro   
        10      clinda  
        0.1     cef     
        0.3     cef
        0.5     cef
        1       metro
        0.1     strep
        0.5     strep
        5       strep
        0.1     vanc
        0.3     vanc
        0.625   vanc
  cdiff - if sample was treated challenged with C. difficile
    logical T(1770), F(303)
  delayed - if sample was allowed extra days to recover from abx treatment
    logical T(455), F(1618)
  preAbx - if sample collected prior to abx treatment
    logical T(154), F(1919)
  recovDays - how many days after stopping abx (metro and amp for 5 day recovery)
    range 1 to 5

only one mouse not given abx
  but is listed as preAbx F for -5
    possible to denote mock abx treatment(?)
question about mouse 600-2D-6 (cef - delivered via water)
  should be pre-antibiotic but is listed as F
  all other mice in cage are preAbx on day -6, except this one
    since this abx was delivered via drinking water, it is likely clerical error, 
    need to write a check script to make sure all mice in each cage all are recorded to have the same treatment

```{r }

library(stringr)


initial_v_final <- hi_lo_no_delay_df  %>% 
  group_by(cage, mouse) %>% 
  summarize(first_day = min(day), last_day = max(day)) %>% ungroup() %>% 
  mutate(initial = paste0(str_pad(cage, 3, pad= '0'), '-', mouse, "D", first_day)) %>% 
  mutate(end = paste0(str_pad(cage, 3, pad= '0'), '-', mouse, "D", last_day)) %>% 
  gather(time_point, group, initial, end) %>% 
  select(group, time_point)

time_nmds_2m <- read.table('data/mothur/abx_time.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.an.unique_list.thetayc.0.03.lt.ave.nmds.2.axes', 
  sep = '\t', header = T)

nmds_2d_plot <- initial_v_final %>% 
  left_join(select(hi_lo_no_delay_df, group, CFU)) %>% 
  left_join(time_nmds_2m) %>% 
  mutate(log_CFU = log10(CFU + 1))

nmds_2d_plot %>% 
  ggplot(aes(x = axis1, y = axis2, color = time_point)) +
    geom_point(aes(size = log_CFU), alpha = 0.5) + 
    theme_bw()


#library(rgl)
#time_nmds_3m <- read.table('data/mothur/abx_time.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.an.unique_list.thetayc.0.03.lt.ave.nmds.3.axes', 
#  sep = '\t', header = T)
#nmds_plot <- initial_v_final %>% 
#  left_join(select(hi_lo_no_delay_df, group, CFU)) %>% 
#  left_join(time_nmds_3m) %>% 
#  mutate(log_CFU = round(log10(CFU + 1) + 1))
#
#par3d(windowRect = c(0, 0, 700, 700))
#with(nmds_plot, 
#  points3d(axis1, axis2, axis3, 
#    xlab = 'axis_1', ylab = 'axis_2', zlab = 'axis_3',
#    width = 1000, size=NULL))
#for(i in c(1,2)){
#  time <- c('initial', 'end')[i]
#  shapes <- c('x','o')[i]
#  pt_color <- c('blue','black')[i]
#  for(sz in c(1:10)){
#    nmds_plot %>% 
#      filter(time_point == time, log_CFU == sz) %>% 
#          with(., text3d(axis1, axis2, axis3, 
#            width = 1000, cex = seq(1,4,0.25)[sz] , color = pt_color, shapes))
#      }
#}
#
#bgplot3d({
#  plot.new()
#  title(main = 'Antibiotic-treated conventional mice challenged with C. difficile', line = 3)
#  mtext(side = 1, 'Points are sized by colonization level',
#     line = 4, cex = 2)
#  legend("topleft", cex = 2, col =c('blue','black'), 
#    pch=c('x','o'), inset=c(0.02),
#    legend = c('Initial', 'Final'))
#  legend("topright", cex = 2,
#    pch='o', pt.cex = c(1.25,1.75,2.25,2.75), inset=c(0.02),
#    legend = c(expression(paste("10"^"2")), expression(paste("10"^"4")),
#      expression(paste("10"^"6")), expression(paste("10"^"8")))
#    )
#})

```